---
title: "Comparing Cluster Output the FORTRAN and R Algorithms Example Data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(LabourMarketAreas)
library(purrr)
library(kableExtra)
```


## load data
```{r, warning=FALSE, include=FALSE}
FORTRAN <- readxl::read_excel(here::here("data/run1410a_output_from_example_data_fortran.xlsx"), 
                                   sheet = "output", 
                                   col_types = "numeric") %>%
  janitor::clean_names() %>%
  rename(ttwa_number = ttw_anumber,
         ttwa_name = ttw_aname)

R <- read_csv(here::here("data/clusterList_output_from_example_data_R.csv"))
```

The aim is to find out how similar the R clusters are to the FORTRAN clusters. 

##
```{r}
# Remove the values which are on the zero list
FORTRAN <- FORTRAN %>% 
  filter(ttwa_number != 0) %>%
  select(zone, ttwa_number) %>%
  arrange(zone)

# Remove the values which should not have formed a cluster (see notes from Mike)
R <- R %>%
  filter(!(community %in% as.vector(setdiff(R$community, FORTRAN$zone)))) %>%
  select(community, cluster) %>%
  arrange(community)
```

```{r create a list of vectors for each cluster}
# create a list of vectors for each cluster 
# summarise clusters as a string of zones in each cluster
FORTRAN_clus <- FORTRAN %>%
   group_by(ttwa_number) %>%
   summarise(
     alltypes = paste(zone, collapse=", ")
   )

# save this vector for use later 
ttwa_number <- as.character(FORTRAN_clus$ttwa_number) 

# split the strings into vectors of strings
FORTRAN_clus <- FORTRAN_clus %>%
  map(function(x) strsplit(FORTRAN_clus$alltypes, ","))

# Tidy up, as we only need one list
FORTRAN_clus <- FORTRAN_clus$ttwa_number

# convert vectors from strings to numeric
FORTRAN_clus <- FORTRAN_clus %>%
  lapply(as.numeric)

#FORTRAN_clus <- setNames(FORTRAN_clus,ttwa_number)


# REPEAT STEPS FOR R OUTPUT
R_clus <- R %>%
   group_by(cluster) %>%
   summarise(
     alltypes = paste(community, collapse=", ")
   )

cluster <- R_clus$cluster
  
R_clus <- R_clus %>%
  map(function(x) strsplit(R_clus$alltypes, ","))

R_clus <- R_clus$cluster

R_clus <- R_clus %>%
  lapply(as.numeric)

R_clus <- setNames(R_clus,cluster)

```



# Similarity Score

One way to compare the outputs is to calculate a similarity score, which assigns a value to a cluster to identify how similar it is to a cluster from the other algorithm. 

```{r}
#jaccard
similarityscore <- function(vectorA, vectorB) {
    intersection = length(intersect(vectorA, vectorB))
    union = length(vectorA) + length(vectorB) - intersection
    return (intersection/union)
}


# Example
vectorA = c(2, 4, 7)
vectorB = c(4, 7, 3)
  
similarityscore(vectorA,vectorB)
```


```{r calculate similarity score, results = 'hide'}
# create function to calculate similarity score
similarityscore <- function(vectorA, vectorB) {
    intersection = length(intersect(vectorA, vectorB))
    union = length(vectorA) + length(vectorB) - intersection
    return (intersection/union)
}

# calculate the similarity scores for the each R_clus vector against each FORTRAN_clus vector to create a matrix
#similarity <- sapply(1:772, function(i) map_dbl(FORTRAN_clus,function(x) similarityscore(R_clus[[i]], x)))
similarity <- sapply(1:785, function(i) map_dbl(R_clus,function(x) similarityscore(FORTRAN_clus[[i]], x)))

# rename the columns
#cluster <- paste0("R_cluster_number_", seq(1,772, by=1))
colnames(similarity, do.NULL = FALSE)
colnames(similarity) <- ttwa_number

# convert to dataframe
similarity <- as_tibble(similarity)
# add column for the ttwa_numbers 
similarity$R_cluster <- cluster
# reorder columns
similarity <- similarity %>%
  select(R_cluster, everything())

# pivot longer so we can compare the similarity scores.
# Filter out all 0 scores
sim_summary <- similarity %>% 
  pivot_longer(!R_cluster, names_to = "F_ttwa_number", values_to = "similarity_score") %>%
  filter(similarity_score > 0)
```


```{r}
sim_summary$comparison_index <- 1:nrow(sim_summary)
sim_summary$R_cluster <- as.numeric(sim_summary$R_cluster)
sim_summary$F_ttwa_number <- as.numeric(sim_summary$F_ttwa_number)

FORTRAN_comp <- left_join(FORTRAN, sim_summary, by = c("ttwa_number" = "F_ttwa_number"))
R_comp <- left_join(R, sim_summary, by = c("cluster" = "R_cluster"))


FORTRAN2 <- FORTRAN %>%
   group_by(ttwa_number) %>%
   summarise(
     zones_F = paste(zone, collapse=", ")
   )

R2 <- R %>%
   group_by(cluster) %>%
   summarise(
     communities_R = paste(community, collapse=", ")
   )

sim_summary2 <- left_join(sim_summary, FORTRAN2, by = c("F_ttwa_number" = "ttwa_number"))
sim_summary2 <- left_join(sim_summary2, R2, by = c("R_cluster" = "cluster"))
View(sim_summary2)

```

## Summary of the similarity scores
```{r, echo=TRUE}
sim_summary %>%
  group_by(similarity_score) %>%
  summarise(n = n()) %>%
  arrange(desc(similarity_score)) %>%
  kbl()
```
## Exact matches

714 clusters have a similarity score of 1, which means the clusters from the R and FORTRAN algorithms match exactly. 

```{r}
sim_summary2 %>% 
  filter(similarity_score == 1) %>%
  select(-c(comparison_index, similarity_score))

```
## Partial matches

```{r}
sim_summary2 %>% 
  filter(similarity_score != 1)

```


```{r}
# The algoritm in R created 772 clusters. 772-714 = 58 
sim_summary %>%
  filter(similarity_score > 0 & similarity_score < 1) %>%
  group_by(R_cluster) %>%
  summarise(n = n()) %>%
  nrow()
# 58 clusters do not match. This makes sense

# The algoritm in FORTRAN created 785 clusters. 785-714 = 71
sim_summary %>%
  filter(similarity_score > 0 & similarity_score < 1) %>%
  group_by(F_ttwa_number) %>%
  summarise(n = n()) %>%
  nrow()
# 71 clusters do not match. This makes sense.

```



